The Second Data Release of the INT Photometric H$\alpha$ Survey of the Northern Galactic Plane (IPHAS DR2)

### 2014-04-13 paper/iphas-dr2.tex
Changes following on from edit by JD

\restylefloat{table}

@@

\begin{document}
\date{Current draft typeset \today}
\pagerange{\pageref{firstpage}--\pageref{lastpage}} \pubyear{2014}

\maketitle

@@

\begin{abstract} % BACKGROUND, OBJECTIVE, METHODS, RESULTS, CONCLUSIONS
The INT/WFC Photometric H$\alpha$ Survey 
of the Northern Galactic Plane (IPHAS)
is a 1800~deg$^2$ imaging survey
covering the Galactic latitude range $|b| < 5^\circ$
in the $r$, $i$ and \ha\ filters 
using the Wide Field Camera (WFC) 
on the 2.5-metre Isaac Newton Telescope (INT) in La Palma.
We present the first quality-controlled and
uniformly-calibrated source catalogue
derived from the survey,
providing single-epoch photometry
for 219 million unique sources
across 92\% of the 1800 square-degree footprint.
The observations were carried out between 2003 and 2013
at a median seeing of 1.1 arcsec
and to mean $5\sigma$-depths of 
21.2 ($r$), 20.0 ($i$) and 20.3 ($\ha$)
in the Vega magnitude system.
We explain the data reduction 
and quality control procedures,
describe and test the new uniform calibration,
and detail the construction of the new 98-column source catalogue.
We show that the global calibration is accurate to
0.03~mag (rms) 
and provide a series of recommended quality criteria
which can be used to select the most reliable data
from the catalogue.
Finally, we demonstrate the ability of the 
catalogue's unique
$(r-\ha,\ r-i)$ diagram to
(i) characterise stellar populations and extinction regimes
towards different Galactic sight-lines
and (ii) select candidate \ha\ emission-line objects.
IPHAS is the first survey to offer comprehensive CCD photometry
of point sources across the Galactic Plane at visible wavelengths,
providing the much-needed counterpart
to recent infrared surveys.
\end{abstract}

@@

\begin{table}
% The caption of tables should be on top and not at the bottom
        \caption{Mean magnitude offsets for objects
                 cross-matched between IPHAS and APASS/SDSS
                 before the uniform re-calibration.
                 Eqns.~\ref{eqn:apass_r}-\ref{eqn:sdss_i} were applied
                 to the APASS/SDSS magnitudes to bring them into the
                 Vega-based IPHAS system prior to computing
                 the offsets.)
                 }
        \label{tbl:offsets_before}
        \begin{center}
                \begin{tabular}{lcc}
                        \toprule
                         {\bf Before re-calibration} & Mean & $\sigma$  \\
                        \midrule
                        $r$ (IPHAS - APASS) & +0.014 & 0.104 \\
                        $i$ (IPHAS - APASS) & +0.007 & 0.108 \\
                        $r$ (IPHAS - SDSS) & +0.016 & 0.088 \\
                        $i$ (IPHAS - SDSS) & +0.010 & 0.089 \\
                        \bottomrule
                \end{tabular}
        \caption{Same as Table~\ref{tbl:offsets_before}
                 but computed after the uniform re-calibration
                 was carried out.}
        \label{tbl:offsets_after}
                \begin{tabular}{lcc}
                        \toprule
                         {\bf After re-calibration} & Mean & $\sigma$ \\
                        \midrule
                        $r$ (IPHAS - APASS) & +0.000 & 0.011\\
                        $i$ (IPHAS - APASS) & +0.000 & 0.011 \\
                        $r$ (IPHAS - SDSS)  & -0.001 & 0.029\\
                        $i$ (IPHAS - SDSS) & -0.002 & 0.032 \\
                        \bottomrule
                \end{tabular}

        \end{center}
\end{table}

@@

Like any other photometric survey,
the majority of the objects in our catalogue
are faint sources observed near the detection limits:
55\% of the entries in the catalogue
are fainter than $r > 20$.
%and 18\% are even fainter than $r > 21$.
The measurements of faint objects
are naturally prone to larger
random and systematic uncertainties:
for example, an inaccurately-subtracted background
will introduce a proportionally larger systematic error
for a faint object.
Most scientific applications will hence require a set of
quality criteria to be enforced for the purpose
of removing lower-quality objects.

@@

Table~\ref{tab:reliable} specifies
a set of minimum quality criteria
which should benefit most applications
which require reliable $(r-i)$ and $(r-\ha)$ colours
(GEERT: the rest of the sentence looks very odd to me -- I can't rephrase
because I'm not sure what you are trying to say) without removing more 
than 80\%
of the sources brighter than $r < 19$.
The listed criteria are designed to 
(i) remove low-S/N sources, 
(ii) remove saturated sources,
and (iii) remove objects for which the 2.3-arcsec diameter
aperture magnitude is inconsistent 
with its alternative 1.2-arcsec diameter measurement.
This last criterion is a proxy
for identifying objects which are affected
by poor background subtraction
or failed source deblending.
A total of 86 out of 219 million sources 
(39\%) satisfy all the criteria listed in Table~\ref{tab:reliable}
and are hereafter referred to as ``reliable''.
For convenience, the catalogue contains a boolean column
named \emph{reliable} that directly flags these objects.

\begin{table*}

\vspace{2cm}
\caption{Recommended minimum quality criteria 
for selecting objects with reliable colours 
from the IPHAS DR2 source catalogue. 
86~million entries in the catalogue (39\%)
satisfy all the criteria listed in this table.
For convenience, these have been flagged in the catalogue
using the column named \emph{reliable}.}
\begin{tabular}{p{8cm}lp{6cm}}
  \hline
  Quality criterion & Rows passed & Description \\
  \hline
  rErr\,$<$\,0.1 {\sc and} iErr\,$<$\,0.1 {\sc and} haErr\,$<$\,0.1 &
  109 million (50\%) &
  Require the photon noise to be less than 
  0.1 mag in all bands (i.e. S/N$>$10).
  This implicitly requires a detection in all three bands.  \\
  $r>13$ {\sc and} $i>12$ {\sc and} \ha\,$>12.5$ {\sc and not} \emph{saturated} &
  158 million (72\%) &
  The brightness must not exceed the nominal saturation limit
  and the peak pixel height must not exceed 55\,000 counts.
  Again, this implicitly requires a detection in all three bands.
  \\
  $|\mathrm{r}- \mathrm{rAperMag1}| 
  < 3\sqrt{\mathrm{rErr}^2 + \mathrm{rAperMag1Err}^2} 
  + 0.03$ &
  176 million (80\%) &
  Require the $r$ magnitude measured 
  in the default 2.3\arcsec\ diameter aperture
  to be consistent with the measurement 
  made in the smaller 1.2\arcsec\ aperture,
  albeit tolerating a 0.03 mag systematic error.
  This will reject sources for which the background
  subtraction or the deblending procedure
  was not performed reliably. \\
  $|\mathrm{i}- \mathrm{iAperMag1}| 
  < 3\sqrt{\mathrm{iErr}^2 + \mathrm{iAperMag1Err}^2} 
  + 0.03$ &
  183 million (84\%) &
  Same as above for $i$. \\
  $|\mathrm{ha}- \mathrm{haAperMag1}| < 
  3\sqrt{\mathrm{haErr}^2 + \mathrm{haAperMag1Err}^2} 
  + 0.03$ &
  158 million (72\%) &
  Same as above for \ha. \\
  \hline
  All of the above (flagged as {\bf\emph{reliable}}) &
  86 million (39\%) & \\
  \hline
\end{tabular}
\label{tab:reliable}

@@

\vspace{2cm}
\end{table*}

For applications which require
an even higher standard of reliability,
a further set of additional quality criteria
are suggested in Table~\ref{tab:veryreliable}.
These criteria are designed to ensure that
(i) the object appeared as a perfect point source,
(ii) the object was not blended with a nearby neighbour,
and (iii) the object was not located near a very bright star.
59 million sources (27\%) satisfy
these additional criteria 
and are hereafter referred to as ``very reliable''.
Again, the catalogue contains a boolean column
named \emph{veryReliable} which flags these objects.

Fig.~\ref{fig:magdist} compares the $r$-band magnitude
distribution for objects with and without the 
\emph{reliable} and \emph{veryReliable} criteria applied. 
We find that 81\% of the sources 
in the magnitude range $13 < r < 19$
are considered \emph{reliable},
which drops to 72\% in the range $19 < r < 20$
and 9\% at $r>20$.
In contrast, only 54\% of the sources 
in the magnitude range $13 < r < 19$
are considered \emph{veryReliable}.
In \S\ref{sec:densities} we will show that these
stricter criteria remove large numbers
of objects at early Galactic longitudes 
where source blending can sometimes affect
more than a quarter of the objects.
The \emph{veryReliable} flag should hence
only be used in applications in which 
completeness is not a concern.

@@

It is easy to see how the quality criteria
may be adapted to be more tolerant.
For example, by raising the allowed photometric uncertainties
from 0.1 mag to 0.2 mag,
42 million candidate sources are added to the 109 
million satisfying the tighter error bound (GEERT - is this what you mean?).

@@

%As we explained earlier, blended sources must be used with caution.
Finally, we warn that blended objects
are more likely than unblended objects to have fallen victim
to source confusion during the band-merging procedure.  It is
important to bear this in mind when appraising stars of seemingly
unusual colours (such as candidate emission line stars) -- if 
blending is flagged, the probability that the unusual colour is 
spurious is greatly increased.
%We have started investigating the use of
%more advanced PSF-fitting routines
%in which sources are measured simultaneously across all bands,
%which may be guided by an external astrometric catalogue
%such as the one that is to be provided by Gaia.

@@

\begin{table}
    \caption{Candidate \ha-emitters towards Sh 2-82.}
    \label{tbl:emitters}
\footnotesize
\resizebox{0.45\textwidth}{!}{
    \begin{tabular}{clccc}
    \toprule
    # & Name [IPHAS2 ...] & $r$ & $i$ & \ha  \\
    \midrule
1 & J192954.40+181026.1& $17.69\pm0.01$ & $16.12\pm0.01$ & $16.19\pm0.01$ \\
2 & J193011.01+182051.2& $18.55\pm0.02$ & $16.95\pm0.02$ & $17.31\pm0.02$ \\
3 & J193021.52+181954.5& $19.72\pm0.05$ & $17.94\pm0.03$ & $18.47\pm0.04$ \\
4 & J193024.45+181938.3& $19.31\pm0.04$ & $17.55\pm0.02$ & $17.99\pm0.03$ \\
5 & J193033.00+181609.3& $18.25\pm0.01$ & $16.91\pm0.01$ & $16.92\pm0.01$ \\
6 & J193042.48+182317.4& $19.96\pm0.03$ & $18.11\pm0.03$ & $18.48\pm0.03$ \\
    \bottomrule
    \end{tabular}}
\end{table}

In previous work, we have shown that the majority of
\ha-emitters seen towards an H{\sc ii} region
are likely to be Classical T Tauri Stars \citep{Barentsen2011a}.
These are young objects
which are thought to show \ha\ in emission
due to the presence of hot, infalling gas
which is accreting onto the star
from a circumstellar disk.
This is likely to be the case for the candidate \ha-emitters
we discovered towards Sh 2-82 as well.
Two of our candidates, x and y in Tab.~\ref{tab:emitters}, have recently 
been identified
as YSOs in the initial investigation of the region by \cite{Yu2012}.
In their study, the authors used 2MASS and Spitzer data
to select likely young stars by selecting objects with
an infrared colour excess, consistent with 
the presence of a heated circumstellar disk.
Although \citeauthor{Yu2012} did not identify
the other four candidate emitters in our sample
as having an infrared excess,
we note that all four objects
are detected in the Spitzer 8.0~\micron\ image
at S/N$>$5,
which is consistent with their
being optically-unveiled stars with a circumstellar disk,
i.e. they are likely to be Class II YSOs exhibiting a mild infrared excess.

@@

A new catalogue of 219 million sources has been
derived from the INT/WFC Photometric \ha\ Survey
of the Northern Galactic Plane.
It is the first catalogue to offer comprehensive CCD photometry
of point sources across the Northern Galactic Plane at visible wavelengths,
taking in the latitude range $|b|<5^\circ$.
The new 98-column catalogue provides single-epoch photometry
across 92\% of the survey area,
and is the first quality-controlled and uniformly calibrated
catalogue to have been constructed from the imaging data.  This
now means that there is H$\alpha$ coverage, accessible online,
of the entire Galactic Plane -- given that the 
southern Plane is already available thanks to the UK
Schmidt H$\alpha$ Survey \citep[SHS,][][Parker05], the last of the 
photographic surveys carried out by that telescope.
%We explain the data reduction and quality control procedures that
%were applied.  We describe and test the global photometric 
%calibration, 
%and detailed the construction of the source catalogue.

@@

The IPHAS survey was carried out 
at the Isaac Newton Telescope (INT).
The INT is operated on the island of La Palma
by the Isaac Newton Group
in the Spanish Observatorio del Roque de los Muchachos
of the Instituto de Astrofisica de Canarias.
We are deeply indebted to the ING staff and students
for their ongoing support of the INT.
All data were processed 
by the Cambridge Astronomical Survey Unit
at the Institute of Astronomy in Cambridge.
The catalogue presented in this work was assembled
at the Centre for Astrophysics Research, University of Hertfordshire, supported by a grant from the Science \& Technology Facilities Council
of the UK (STFC, ref ST/J001335/1).

@@

\small
\begin{longtable}{lllp{10cm}}
\caption{\label{tab:columns} 
Definition of columns in the IPHAS DR2 source catalogue.
} \\
\hline
Column & Type & Unit & Description \\
\hline
\endfirsthead

\multicolumn{3}{c}%
{{\bfseries \tablename\ \thetable{} -- continued}} \\
\hline
Column & Type & Unit & Description \\
\hline
\endhead

>>>>

@@

\begin{document}
\date{Current draft typeset \today}
\pagerange{\pageref{firstpage}--\pageref{lastpage}} \pubyear{2014}
\maketitle

@@

\begin{abstract} % BACKGROUND, OBJECTIVE, METHODS, RESULTS, CONCLUSIONS
The INT/WFC Photometric H$\alpha$ Survey 
of the Northern Galactic Plane (IPHAS)
is a 1800~deg$^2$ imaging survey
covering Galactic longitudes $l=30^\circ$ to $215^\circ$
and latitudes $b = -5^\circ$ to $+5^\circ$
in the $r$, $i$ and \ha\ filters 
using the Wide Field Camera (WFC) 
on the 2.5-metre Isaac Newton Telescope (INT) in La Palma.
We present the first quality-controlled and
uniformly-calibrated source catalogue
derived from the survey,
providing single-epoch photometry
for 219 million unique sources
across 92\% of the footprint.
The observations were carried out between 2003 and 2013
at a median seeing of 1.1 arcsec
and to a mean $5\sigma$-depth of 
21.2 ($r$), 20.0 ($i$) and 20.3 ($\ha$)
in the Vega magnitude system.
We explain the data reduction 
and quality control procedures,
describe and test the new uniform calibration,
and detail the construction of the new 98-column source catalogue.
We show that the global calibration is accurate to
0.03~mag (rms) 
and recommend a series of quality criteria
to select the most reliable data
from the catalogue.
Finally, we demonstrate the ability of the 
catalogue's unique
$(r-\ha,\ r-i)$ diagram to
(i) characterise stellar populations and extinction regimes
towards different Galactic sight-lines
and (ii) select candidate \ha\ emission-line objects.
IPHAS is the first survey to offer comprehensive CCD photometry
of point sources across the Galactic Plane at visible wavelengths,
providing the much-needed counterpart
to recent infrared surveys.
\end{abstract}

@@

\begin{table}
    \caption{Mean magnitude offsets for objects
             cross-matched between IPHAS and APASS/SDSS
             before the uniform re-calibration.
             Eqns.~\ref{eqn:apass_r}-\ref{eqn:sdss_i} were applied
             to the APASS/SDSS magnitudes to bring them into the
             Vega-based IPHAS system prior to computing
             the offsets.)
             }
    \label{tbl:offsets_before}
    \begin{center}
        \begin{tabular}{lcc}
            \toprule
            {\bf Before re-calibration} & Mean & $\sigma$  \\
            \midrule
            $r$ (IPHAS - APASS) & +0.014 & 0.104 \\
            $i$ (IPHAS - APASS) & +0.007 & 0.108 \\
            $r$ (IPHAS - SDSS) & +0.016 & 0.088 \\
            $i$ (IPHAS - SDSS) & +0.010 & 0.089 \\
            \bottomrule
       \end{tabular}
       \caption{Same as Table~\ref{tbl:offsets_before}
                but computed after the uniform re-calibration
                 was carried out.}
        \label{tbl:offsets_after}
            \begin{tabular}{lcc}
                \toprule
                {\bf After re-calibration} & Mean & $\sigma$ \\
                \midrule
                $r$ (IPHAS - APASS) & +0.000 & 0.011\\
                $i$ (IPHAS - APASS) & +0.000 & 0.011 \\
                $r$ (IPHAS - SDSS)  & -0.001 & 0.029\\
                $i$ (IPHAS - SDSS) & -0.002 & 0.032 \\
                \bottomrule
            \end{tabular}
        \end{center}
\end{table}

@@

Like any other photometric survey,
the majority of the objects in our catalogue
are faint sources observed near the detection limits;
55\% of the entries in the catalogue
are fainter than $r > 20$.
%and 18\% are even fainter than $r > 21$.
The measurements of faint objects
are naturally prone to larger
random and systematic uncertainties:
for example, an inaccurately-subtracted background
will introduce a proportionally larger systematic error
for a faint object.
Most scientific applications will hence require a set of
quality criteria to be enforced for the purpose
of removing lower-quality objects.

@@

Table~\ref{tab:reliable} specifies
a set of minimum quality criteria
which should benefit most applications
which desire reliable $(r-i)$ and $(r-\ha)$ colours
as well as completeness.
The listed criteria are designed to 
(i) remove low-S/N sources, 
(ii) remove saturated sources,
and (iii) remove objects for which the 2.3-arcsec diameter
aperture magnitude is inconsistent 
with its alternative 1.2-arcsec diameter measurement.
This last criterion is a proxy
for identifying objects which are affected
by poor background subtraction
or failed source deblending.
A total of 86 out of 219 million sources 
(39\%) satisfy all the criteria listed in Table~\ref{tab:reliable}
and are hereafter referred to as ``reliable''.
For convenience, the catalogue contains a boolean column
named \emph{reliable} that directly flags these objects.

\begin{table*}
\vspace{2cm}
\caption{Recommended minimum quality criteria 
for selecting objects with reliable colours 
from the IPHAS DR2 source catalogue. 
86~million entries in the catalogue (39\%)
satisfy all the criteria listed in this table.
For convenience, these have been flagged in the catalogue
using the column named \emph{reliable}.}
\begin{tabular}{p{8cm}lp{6cm}}
  \hline
  Quality criterion & Rows passed & Description \\
  \hline
  rErr\,$<$\,0.1 {\sc and} iErr\,$<$\,0.1 {\sc and} haErr\,$<$\,0.1 &
  109 million (50\%) &
  Require the photon noise to be less than 
  0.1 mag in all bands (i.e. S/N$>$10).
  This implicitly requires a detection in all three bands.  \\
  $r>13$ {\sc and} $i>12$ {\sc and} \ha\,$>12.5$ {\sc and not} \emph{saturated} &
  158 million (72\%) &
  The brightness must not exceed the nominal saturation limit
  and the peak pixel height must not exceed 55\,000 counts.
  Again, this implicitly requires a detection in all three bands.
  \\
  $|\mathrm{r}- \mathrm{rAperMag1}| 
  < 3\sqrt{\mathrm{rErr}^2 + \mathrm{rAperMag1Err}^2} 
  + 0.03$ &
  176 million (80\%) &
  Require the $r$ magnitude measured 
  in the default 2.3\arcsec\ diameter aperture
  to be consistent with the measurement 
  made in the smaller 1.2\arcsec\ aperture,
  albeit tolerating a 0.03 mag systematic error.
  This will reject sources for which the background
  subtraction or the deblending procedure
  was not performed reliably. \\
  $|\mathrm{i}- \mathrm{iAperMag1}| 
  < 3\sqrt{\mathrm{iErr}^2 + \mathrm{iAperMag1Err}^2} 
  + 0.03$ &
  183 million (84\%) &
  Same as above for $i$. \\
  $|\mathrm{ha}- \mathrm{haAperMag1}| < 
  3\sqrt{\mathrm{haErr}^2 + \mathrm{haAperMag1Err}^2} 
  + 0.03$ &
  158 million (72\%) &
  Same as above for \ha. \\
  \hline
  All of the above (flagged as {\bf\emph{reliable}}) &
  86 million (39\%) & \\
  \hline
\end{tabular}
\label{tab:reliable}

@@

\vspace{1cm}
\end{table*}

For applications which require
a higher standard of reliability at the expense of completeness,
a further set of additional quality criteria
are suggested in Table~\ref{tab:veryreliable}.
These criteria are designed to ensure that
(i) the object appeared as a perfect point source,
(ii) the object was not blended with a nearby neighbour,
and (iii) the object was not located near a very bright star.
59 million sources (27\%) satisfy
these stricter criteria 
and are hereafter referred to as ``very reliable''.
Again, the catalogue contains a boolean column
named \emph{veryReliable} which flags these objects.

Fig.~\ref{fig:magdist} compares the $r$-band magnitude
distributions for \emph{reliable}, \emph{veryReliable}
and unfiltered objects.
We find that 81\% of the sources 
are considered \emph{reliable}
and 54\% are \emph{veryReliable}
in the magnitude range $13 < r < 19$.
We will explain below that the
\emph{veryReliable} category is least complete
at low Galactic longitudes,
where source blending can affect
up to a quarter of the objects.
The \emph{veryReliable} flag should hence
only be used in applications in which 
completeness is not a concern.

@@

It is easy to see how the quality criteria
may be adapted to be more tolerant.
For example, by raising the allowed photometric uncertainties
from 0.1~mag to 0.2~mag,
42 million candidate sources would be added to the 109 
million satisfying the tighter error bound.
Our choice to adopt 0.1~mag as the cut-off uncertainty
for the \emph{reliable} category is a pragmatic trade-off
which we found to suit many science applications,
but certainly not all. Users of the data are encouraged 
to revise the quality criteria according to their needs.

@@

Finally, we warn that blended objects
are more likely than unblended objects to have fallen victim
to source confusion during the band-merging procedure.  It is
important to bear this in mind when appraising stars of seemingly
unusual colours (such as candidate emission line stars) -- if 
blending is flagged, the probability that the unusual colour is 
spurious is greatly increased.
%We have started investigating the use of
%more advanced PSF-fitting routines
%in which sources are measured simultaneously across all bands,
%which may be guided by an external astrometric catalogue
%such as the one that is to be provided by Gaia.

@@

\begin{table}
    \caption{Candidate \ha-emitters towards Sh 2-82.}
    \label{tbl:emitters}
    \footnotesize
    \resizebox{0.45\textwidth}{!}{
        \begin{tabular}{clccc}
        \toprule
        \# & Name [IPHAS2 ...] & $r$ & $i$ & \ha  \\
        \midrule
        1 & J192954.40+181026.1& $17.69\pm0.01$ & $16.12\pm0.01$ & $16.19\pm0.01$ \\
        2 & J193011.01+182051.2& $18.55\pm0.02$ & $16.95\pm0.02$ & $17.31\pm0.02$ \\
        3 & J193021.52+181954.5& $19.72\pm0.05$ & $17.94\pm0.03$ & $18.47\pm0.04$ \\
        4 & J193024.45+181938.3& $19.31\pm0.04$ & $17.55\pm0.02$ & $17.99\pm0.03$ \\
        5 & J193033.00+181609.3& $18.25\pm0.01$ & $16.91\pm0.01$ & $16.92\pm0.01$ \\
        6 & J193042.48+182317.4& $19.96\pm0.03$ & $18.11\pm0.03$ & $18.48\pm0.03$ \\
        \bottomrule
        \end{tabular}
    }
\end{table}

In previous work, we have shown that the majority of
\ha-emitters that are unveiled by IPHAS towards an H{\sc ii}
are likely to be Classical T Tauri Stars \citep{Barentsen2011a}.
These are young objects
which are thought to show \ha\ in emission
due to the presence of hot, infalling gas
which is accreting onto the star
from a circumstellar disk.
This is likely to be the case for the candidate \ha-emitters
we discovered towards Sh 2-82 as well.
Two of our candidates, \#1 and \#4 in Tab.~\ref{tbl:emitters},
have previously been identified
as candidate YSOs by \cite{Robitaille2008}
and \cite{Yu2012}, respectively.
In their studies, the authors used Spitzer data
to select intrinsically red objects,
consistent with the presence of a circumstellar disk.
Although the other four candidate emitters in our sample
have not previously appeared in the literature,
we note that all four are detected
in the Spitzer 8.0~\micron\ image at S/N$>$5,
which is consistent with being optically-unveiled stars with a circumstellar disk,
i.e. they are likely to be Class II YSOs exhibiting a mild infrared excess.

@@

A new catalogue of 219 million sources has been
derived from the INT/WFC Photometric \ha\ Survey
of the Northern Galactic Plane.
It is the first catalogue to offer comprehensive CCD photometry
of point sources across the Northern Galactic Plane at visible wavelengths,
taking in the latitude range $|b|<5^\circ$.
The new 98-column catalogue provides single-epoch photometry
across 92\% of the survey area,
and is the first quality-controlled and uniformly calibrated
catalogue to have been constructed from the imaging data.  This
now means that there is H$\alpha$ coverage, accessible online,
of the entire Galactic Plane -- given that the 
southern Plane is already available thanks to the UK
Schmidt H$\alpha$ Survey \citep[SHS,][]{Parker2005}, the last of the 
photographic surveys carried out by that telescope.

@@

The INT is operated on the island of La Palma
by the Isaac Newton Group
in the Spanish Observatorio del Roque de los Muchachos
of the Instituto de Astrofisica de Canarias.
We are deeply indebted to the ING staff and students
for their ongoing support of the INT.
All data were processed 
by the Cambridge Astronomical Survey Unit
at the Institute of Astronomy in Cambridge.
The catalogue presented in this work was assembled
at the Centre for Astrophysics Research, University of Hertfordshire, supported by a grant from the Science \& Technology Facilities Council
of the UK (STFC, ref ST/J001335/1).

@@

\small
\begin{longtable}{llllp{10cm}}
\caption{\label{tab:columns} 
Definition of columns in the IPHAS DR2 source catalogue.
} \\
\hline
\# & Column & Type & Unit & Description \\
\hline
\endfirsthead

\multicolumn{3}{c}%
{{\bfseries \tablename\ \thetable{} -- continued}} \\
\hline
\# & Column & Type & Unit & Description \\
\hline
\endhead
